---
title: "DHiam_BikeShare_Project 2"
output:
  html_document: default
  pdf_document: default
---

**Set Working Directory and Load files**

```{r echo=FALSE}
getwd()
setwd("C:/Users/e5104965/Dropbox/RMIT_Training/R_Bikeshare")
list.files()

ny = read.csv('new-york-city.csv')
wash = read.csv('washington.csv')
chi = read.csv('chicago.csv')
```

**Upload Packages required**

```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
```

**Bind all 3 datasets into single dataset to get results for overall findings**

```{r}
Bind_ALL=bind_rows(ny, chi, wash)
dim(Bind_ALL)
```

**NOTE**
Each dataset has 300000 observations and 9 variable expcept for Washington who does not have gender or Birth year data and these are allocated NA. While there are warnings they are acceptable. 


**Popular times of travel (i.e., occurs most often in the start time)**

*QUESTION ONE: What is the most common month?*
```{R}
Bind_ALL$month <- month(dmy_hm(Bind_ALL$Start.Time), label= TRUE, abbr=FALSE) 
```
This code above pulls the month out of the Start.Time colunmn and allows us to then count the frequency

```{r echo= TRUE, results = 'asis'}
library(knitr)
month <- Bind_ALL %>%  
group_by(month) %>%
  summarise(frequency = n())
kable(month)
```

 
VISUALISE Q1

```{r}
Graph=as.data.frame(month)
ggplot(aes(x=month, y=frequency), data=month) +
  geom_col()+
  theme_classic()+
  labs(y="Frequency of hire", 
       x="Month", 
       title="Frequency of bike hire by month", 
       caption = "Source: Motivate (New York, Washington, Chicago)")
```

**Answer: According to the graph and the numeric summary June is the most common month for bike hires.**


*What is the most common day of the week*

```{r}
Bind_ALL$day <-wday(dmy_hm(Bind_ALL$Start.Time), label=TRUE, abbr = FALSE)
```


```{r echo= TRUE, results = 'asis'}
library(knitr)
day <- Bind_ALL %>%
  group_by(day) %>%
  summarise(frequency = n())
kable(day)
```

Visualise Q2
```{r}
Graph=as.data.frame(day)
dim(day)
ggplot(aes(x=day, y=frequency), data=day) +
  geom_col()+
  theme_classic()+
  labs(y="Frequency of hire", 
       x="Day of week", 
       title="Frequency of bike hire by day", 
       caption = "Source: Motivate (New York, Washington, Chicago)")
```

**Answer: According to the graph and numeric summary Wednesday is the most common month for bike hires.** 


*QUESTION 3: What is the most common hour of day?*

```{r}
Bind_ALL$hour <- hour(dmy_hm(Bind_ALL$Start.Time))
```

```{r echo= TRUE, results = 'asis'}
hour <- Bind_ALL %>%
  group_by(hour) %>%
  summarise(frequency = n())
kable(hour)
```

Visualise Q3
```{r}
Graph=as.data.frame(hour)
ggplot(aes(x=hour, y=frequency), data=hour) +
  geom_col()+
  theme_classic()+
  scale_x_continuous(breaks = 0:23)+
  labs(y="Frequency of hire", 
       x="Hour of the day (24 hour clock format)", 
       title="Frequency of bike hire by hour", 
       caption = "Source: Motivate (New York, Washington, Chicago)")
```

**According to the graph and numeric summary (labelled hour) 8am is the most common hour for bike hires.** 

